validateHOT - an R package for holdout task validation and market simulations

### 2024-03-04 paper/paper.Rmd
deleted references and also changed affiliation - 240304

---
title: 'validateHOT - an R package for holdout task validation and market simulations'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Benjamin Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1, 2
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters
date: 02 January
year: 2024
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
header-includes: 
 - \usepackage{float}
 - \usepackage{fancyhdr}
---

@@

Preference measurement techniques' (e.g., (A)CBC or MaxDiff) aim is to predict behavior [@green1990]. Hence, it is essential to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well. The easiest way for testing validity is by including validation tasks [e.g., @rao2014; @Orme2015], which are fixed tasks (i.e., same across participants) and not used for estimating the part-worth utilities (raw logit utilities) in hierarchical Bayes (HB) estimation. Despite their importance, practitioners don't always include them [@yang2018]. This is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

@@

Other packages provide functions to calculate validation metrics, however, these are not always specified for individual part-worth utilities. Metrics [@Metrics], for example, provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to get the output of, for example, Sawtooth Software or ChoiceModelR [@ChoiceModelR] into the right format, the user needs some data wrangling. The package conjoint [@conjoint] provides functions that are most similar to validateHOT's ones, but no validation functions are included and the package focuses on classical conjoint analysis, thus it is limited when applying more common conjoint methods. support.BWS [@support.BWS] only covers best-worst scaling case 1 (i.e., MaxDiff). logitr [@logitr] provides market simulations tools, however, no validation metrics such as mean hit probability [@voleti2017] or hit rate [@netzer2011]. \autoref{comparison} shows a comparison of validateHOT's functions with current R packages. To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

@@

validateHOT's functions can be categorized into four main components, see \autoref{tab:table1}. To bring the data into the right format for some functions, the \texttt{\color{purple}createHOT()} function can be applied, which creates each alternatives' total utility by applying the additive utility model [@rao2014, p. 82].

@@

After running the HB estimation [@allenby1995; @lenk1996], the **raw** utility scores have to be exported and read into an *R* data frame. 
Assuming you included a validation task with seven alternatives plus the no-buy alternative (\texttt{\color{purple}none}). To create this validation task in *R*, we use the \texttt{\color{purple}createHOT()} function.

@@

To get the relevant validation metrics that are reported in conjoint studies, for example, hit rate [e.g., @ding2005] or mean hit probability [mhp, @voleti2017], we provide the data, the alternatives in the validation task (\texttt{\color{purple}opts}), and the actual choice (\texttt{\color{purple}choice}). The function can be implemented using the tidyverse [@tidyverse] logic.

>>>>

---
title: 'validateHOT - an R package for holdout task validation and market simulations'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Benjamin Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 1    
affiliations:
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 1
citation_author: Schramm & Lichters
date: 04 March
year: 2024
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
header-includes: 
 - \usepackage{float}
 - \usepackage{fancyhdr}
---

@@

Preference measurement techniques' (e.g., (A)CBC or MaxDiff) aim is to predict behavior [@green1990]. Hence, it is essential to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well. The easiest way for testing validity is by including validation tasks [e.g., @Orme2015], which are fixed tasks (i.e., same across participants) and not used for estimating the part-worth utilities (raw logit utilities) in hierarchical Bayes (HB) estimation. Despite their importance, practitioners don't always include them [@yang2018]. This is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

@@

Other packages provide functions to calculate validation metrics, however, these are not always specified for individual part-worth utilities. Metrics [@Metrics], for example, provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to get the output of, for example, Sawtooth Software or ChoiceModelR [@ChoiceModelR] into the right format, the user needs some data wrangling. The package conjoint [@conjoint] provides functions that are most similar to validateHOT's ones, but no validation functions are included and the package focuses on classical conjoint analysis, thus it is limited when applying more common conjoint methods. logitr [@logitr] provides market simulations tools, however, no validation metrics such as mean hit probability or hit rate. \autoref{comparison} shows a comparison of validateHOT's functions with current R packages. To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

@@

validateHOT's functions can be categorized into four main components, see \autoref{tab:table1}. To bring the data into the right format for some functions, the \texttt{\color{purple}createHOT()} function can be applied, which creates each alternatives' total utility by applying the additive utility model.

@@

After running the HB estimation [@allenby1995], the **raw** utility scores have to be exported and read into an *R* data frame. 
Assuming you included a validation task with seven alternatives plus the no-buy alternative (\texttt{\color{purple}none}). To create this validation task in *R*, we use the \texttt{\color{purple}createHOT()} function.

@@

To get the relevant validation metrics that are reported in conjoint studies, for example, hit rate or mean hit probability, we provide the data, the alternatives in the validation task (\texttt{\color{purple}opts}), and the actual choice (\texttt{\color{purple}choice}). The function can be implemented using the tidyverse [@tidyverse] logic.
